---
title: "DATA_607_Week_5_Assignment"
author: "Md Jalal Uddin"
date: "October 1, 2016"
output: pdf_document
---

```{r}
# Load packages 

library(dplyr)
library(tidyr)
library(stringr)
library(knitr)

```


```{r}
flight_delay <- read.csv("C:/Users/sql_ent_svc/Google Drive/DATA_607/Week_5_Assignment/Arrival_delay.csv", header = TRUE, sep = "," , stringsAsFactors = FALSE)
flight_delay
#read the csv file using read.csv function and assign it into flight_delay variable. 

```

writing the flight name "Alaska" and "AM WEST" to the empty space
```{r}
names(flight_delay)[names(flight_delay) == "X"] <- "Airline"
names(flight_delay)[names(flight_delay) == "X.1"] <- "Arrival"
flight_delay[2, 1] <- "Alaska"
flight_delay[4, 1] <- "AM WEST"
flight_delay

```

reshape the data with all corresponded values  by using gather funtion
```{r}
library(tidyr)
flight_delay <- gather(flight_delay, "City", "Count", 3:7)
flight_delay

#Source:http://garrettgman.github.io/tidying/

```

Reshape the flight_delay data in regards to key colum "Arrival" and minimizes the duplication of the data
```{r}

flight_delay <- spread(flight_delay, "Arrival", Count)
flight_delay

#Source:http://garrettgman.github.io/tidying/

```

by using mutate function I created total column which is the addition of number of delay and ontime flights and another new delayPerchetage column by dividing the delayed column by Total and multiplied by 100
```{r}
library(dplyr)
flight_delay <- mutate(flight_delay, Total = delayed + ontime, delayPercentage = delayed / Total * 100)
flight_delay <- arrange(flight_delay, delayPercentage)
flight_delay

```

Creating barchart using ggplot graphical function
```{r}
library(ggplot2)
ggplot(flight_delay,aes(x=City,y=delayPercentage,fill=factor(Airline)))+
    geom_bar(stat="identity",position="dodge")

#Source:http://docs.ggplot2.org/0.9.3.1/geom_bar.html
```


Creating a piechart 
```{r}
library(ggplot2)
delay_per <- c(5.15, 7.89, 8.62, 11.09, 14.21, 14.42, 14.50,16.85, 23.28,28,73)
City <- c("Phoenix(Alaska)", "Phoenix(AM WEST)", "San.Diego(Alaska)", "Los.Angeles(Alaska)", "Seattle(Alaska)", "Los.Angeles(AM WEST)", "San.Diego(AM WEST)", "San.Francisco(Alaska)", "Seattle(AM WEST)", "San.Francisco(AM WEST)" )
City <- paste(City, "%", sep = "")  # ad % to labels 
pie(delay_per, labels = City, col = rainbow(length(City)), main = "Pie Chart of Delay Flights")

#Source:http://www.statmethods.net/graphs/pie.html

```



